The MACE is essentially unidimensional: Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' scale

### 2022-08-08 manuscript/main.tex
manuscipt edits (2022-08-07b)

Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poorer physical health \citep{wegman2009meta}. Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}. Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

@@

An overall shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance. One possibility is the replication sample had simply experienced more adversity than the original sample. A second is that the differences in the demographic composition of the two samples (e.g. gender, age) explain the differences in endorsement rates. To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's endorsement response on an item is regressed against their rest score (i.e. their total score for all other items), sample membership, gender, and age.

@@

To evaluate the reliability of the MACE total and subscale scores, we computed several model-based reliability indices using the factor loading estimates from the bifactor models. First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e. both group and specific factors). Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}). Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor. The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores. Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score. When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor even if the data are multidimensional \citep{rodriguez2016applying}. So too values of $\omega_s$ approaching 1 indicate the specific factor, and not the general factor, is the primary source of reliable variance in a subscale score. \cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable. In contrast, values of $\omega_s$ smaller than 0.50 preclude the interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}.

To assess the unidimensionality of the MACE, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}). The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale; it therefore measures the contribution of the general factor relative to the specific factors. As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble those that would be obtained by a one-dimensional model. ECV values greater than 0.70 often indicate a scale is essentially unidimensional \citep{rodriguez2016applying}. In turn, the PUC quantifies how many inter-item correlations are accounted for only by the general factor and is an important moderator of the ECV. As the PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure. When PUC > 0.80, there is a low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}. As an additional check, we calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading the bifactor, divided by the general factor loading in the bifactor. Parameter bias less than 10–15\% is acceptable \citep{muthen1987structural}.

Finally, we calculated the H index to assess the construct replicability of each factor \citep{hancock2001rethinking}. The H index is an estimate of how well a set of items represents a latent variable. H values closer to 1 indicate a well-defined latent variable with factor loadings more likely to be stable across studies. In contrast, small H values suggest a poorly defined latent variable where factor loadings are liable to change across studies. Because we are primarily focused on measuring the reliability of MACE total and subscale scores, we report the H values below but do not discuss them further.

@@

In addition to confirmatory factor analysis, we also we also performed exploratory factor analysis (EFA) on the response data from each sample separately. The goal of the EFA was to observe what factor structure would emerge with no \emph{a priori} restrictions. We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models. We were particularly interested in seeing if a two-factor solution would produce the factors from model 3 (i.e. threat and deprivation); and if a three-factor solution would produce the three factors from model 4 (i.e. maltreatment by adults, peer victimization, reverse-scoring).

EFA was performed using the \textit{lavaan} (v0.6.11; \citealt{lavaan}) package available in R. Factors were extracted using the oblique geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings. We used two rotation criteria in order to examine the generalizability of the factor solutions across multiple rotations. We again relied on traditional fit indices (i.e. RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory models to the response data of both samples.

@@

@@

The item discrimination parameters of the unidimensional model were, on average, large for both the original sample (mean $\alpha$ = 1.357, 95\% HDI = 1.295 -- 1.414) and replication sample (mean $\alpha$ = 1.349, 95\% HDI = 1.272 -- 1.422). The standardized factor loadings were correspondingly large in magnitude. In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621). Most items had loadings greater than 0.50 (original: 80.2\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment. The rank correlation of factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement on the relative loading of items onto the common factor across samples.

To evaluate the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric correlation matrix for each sample. In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they are 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56. The first-to-second eigenvalue ratios are 3.84 and 3.02, respectively, just above the 3-to-1 criterion often cited in determining whether item response data are essentially unidimensional \citep{embretson2013item}.

In sum, the fits of the one-factor models provide partial evidence in support of a unidimensional account of the MACE. The items exhibited uniformly moderate-to-large loadings onto the common factor. Moreover, the eigenvalues of the polychoric correlation matrix suggest one principal factor. However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be additional dimensions to the MACE. We turn next to the multidimensional confirmatory models to explore these possibilities.

@@

Next we inspect the bifactor model with 10 specific factors, i.e. one per MACE subscale (model 2). In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864, with an average value of 0.575 (95\% HDI = 0.556 -- 0.593). In the replication sample, the general factor loadings ranged between 0.299 and 0.827, with an average value of 0.580 (95\% HDI = 0.560 -- 0.601). The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), suggesting little bias in the factor loadings of the unidimensional models due to multidimensionality.

By contrast the standardized loadings on the specific factors were much smaller in magnitude, seldom exceeding loadings on the general factor. In the original dataset, the specific factor loadings ranged 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341). In the replication dataset, the range was between 0.043 and 0.698, with an average value of 0.343 (95\% HDI = 0.315 -- 0.369). Only items from the peer victimization subscales (i.e. peer verbal \& physical abuse) had specific factor loadings of roughly equal or greater magnitude to those on the general factor.

@@

Given these, we found that the MACE overall severity score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965). Further, the $\omega_h$ value for both samples was 0.924 indicating that 96.0\% and 95.8\% of the reliable variance in the overall severity score is attributable to the general maltreatment factor. (This is not in itself so surprising given the large number of subscales and relative paucity of items per subscale.) Notably, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively. In conjunction with a large PUC value of 0.912, these results provide clear evidence for a dominant general maltreatment factor in the MACE.

The reliability of the MACE subscale scores were more variable in comparison. The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample, with smaller values for subscales with fewer items. Crucially, the $\omega_s$ values were mostly small for both the original sample (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and replication sample (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565). That is, the MACE subscale scores contain almost no reliable variance independent of the general maltreatment factor. Therefore they should not be interpreted as reflecting specific forms of maltreatment. The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores (original = 70.7\%; replication = 66.2\%) was independent of general maltreatment.

The results of the bifactor model were unequivocal. Responses on the MACE are characterized by a dominant general maltreatment factor with strong loadings from the majority of items. In stark contrast, the specific factors corresponding to each MACE subscale are weak and characterized mostly by small factor loadings. Whereas the MACE overall severity score is reliable and a valid measure of general maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to specific types of abuse. The question remains, however, if there are other valid subscale scores to be derived from the MACE in addition to the overall severity score.

\subsubsection{Model 3: Bifactor S-1 model with one specific factor (deprivation)}

We turn our attention next to the threat-deprivation bifactor S-1 model (model 3). The loadings on the general factor largely resembled those of the bifactor model. In the original sample, the loadings on the general maltreatment factor ranged between 0.150 and 0.864, with an average value of 0.577 (95\% HDI = 0.558 -- 0.594). In the replication sample, the general factor loadings ranged between 0.395 and 0.831, with an average value of 0.580 (95\% HDI = 0.558 -- 0.601).

The loadings on the specific (neglect) factor were smaller on average. In the original sample, the average specific loading was 0.410 (95\% HDI = 0.373 -- 0.454); in the replication sample, it was 0.454 (95\% HDI = 0.409 -- 0.499). The specific loadings exhibited a notable pattern where only the six reverse-scored items exhibited moderate loadings; the remaining four item exhibited negligible factor loadings. In other words, the ``neglect'' factor appears instead to reflect a reverse-scoring methods factor.

As observed with the bifactor model, the overall severity score under the threat- deprivation bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959). Unsurprisingly, the $\omega_h$ values indicated the majority of reliable variance in severity score is attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922). The reliability of the ``neglect'' subscale score was similarly good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921). The $\omega_s$ values, however, indicated the subscale score contained less than half reliable variance independent of the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).

In summary then, the results do not support the use of threat and deprivation subscale scores as derived from the abuse and neglect items on the MACE. The minority of reliable variance unique from general maltreatment prevents the deprivation score from being interpreted as such. Furthermore, the items comprising deprivation subscale seem to be confounded by a reverse-scoring methods factor.

@@

Finally, we inspect the bifactor S-1 model with two specific factors measuring peer victimization and reverse-scoring (model 4). In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864, with an average value of 0.563 (95\% HDI = 0.545 -- 0.580). In the replication sample, the general factor loadings ranged between 0.298 and 0.852, with an average value of 0.572 (95\% HDI = 0.550 -- 0.593). Interestingly, the loadings on the specific factors were of approximately equal magnitude. In the original dataset, the specific factor loadings ranged 0.327 to 0.789 with an average value of 0.550 (95\% HDI = 0.522 -- 0.576). In the replication dataset, the range was between 0.183 and 0.684, with an average value of 0.526 (95\% HDI = 0.492 -- 0.556).

As with the other bifactor models, the overall severity score under this bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961). Again the majority of reliable variance in severity score was attributable to the general factor. The peer victimization subscale score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868). Notably, the corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample. That is, slightly more than half of the variance in peer victimization scores (original: 64.0\%; replication: 56.8\%) is unique from the general maltreatment factor. These results suggest there exists a secondary factor in the MACE, peer victimization, independent of general maltreatment (albeit marginally so).

Briefly, we note that the subscale scores formed by the reverse-scored items was also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915). Moreover, the corresponding $omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498). As such, these scores are also have majority reliable variance independent of general maltreatment. It is difficult to interpret these scores, however, as it is unclear if they primarily reflect neglect experiences, a wording effect, or a scoring-induced methods artifact.

@@

The goodness-of-fit of the exploratory factor models to the response data of both samples are summarized in Table \ref{tab:efa_diagnostics}. Briefly, the fit indices indicated that all models provided an adequate fit to the data. The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are summarized in Figure \ref{fig:efa_cf}. (The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, there was a primary adult maltreatment factor and a secondary peer victimization factor. In contrast, the replication sample produced a primary adult/peer maltreatment factor with a secondary reverse-scoring factor.

The three-factor solution for each sample were similar. The analysis produced a primary adult maltreatment factor with secondary peer victimization and reverse-scoring factors. (To note items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, warranting some caution in interpreting the replicability of factor structures.) With the oblique rotation, we observed only modest inter-factor correlations (original: mean $r$ = 0.371; replication: mean $r$ = 0.269). To conclude, the structure of both the two- and three-factor exploratory factor solutions resembled the structure of the bifactor S-1 model with two specific factors (model 4), providing further support for this model of responses on the MACE.

@@

The MACE is a promising alternative measure of childhood maltreatment \citep{georgieva2022systematic}. However, there has been little study of its structural validity \citep{saini2019systematic}, leaving unanswered questions about the scale's underlying factor structure and the validity of its total/subscale scores. Here we investigated the factor structure of the MACE using both confirmatory and exploratory item response modeling. The fits of a bifactor model (model 2) to the response data from both samples revealed a dominant general factor with uniformly strong loadings from the majority of items, and substantially weaker loadings on the specific subscale factors. The general maltreatment factor explained approximately 70\% of the reliable variance in responses across samples. When comparing the loadings on the general factor to those from a unidimensonal model, the relative parameter bias was marginal. Together, these results support the conclusion that the MACE is essentially unidimensional.

We also calculated model-based reliability indices of the MACE total and subscale scores using the bifactor model fits. We found that the MACE total score has excellent reliability and is a valid measure of general maltreatment. In contrast, we found that the MACE subscale scores were not valid measures of specific forms of maltreatment; in all but one example, the majority of reliable variance in subscale scores was attributable to the general factor, thereby precluding meaningful interpretation of the MACE subscale scores as such. The same was true of the deprivation subscale scores, derived from the neglect items on the MACE, which also possessed little reliable variance independent of the general factor and was contaminated by a reverse-scoring methods factor. In sum, whereas the MACE total score is reliable and valid, subscale scores derived from a subset of items are seldom interpretable as such and instead reflect general maltreatment.

Of the four confirmatory factor models we tested, the best-fitting and most parsimonious was a bifactor S-1 model with dominant general maltreatment factor and two smaller specific factors reflecting peer victimization and reverse-scoring (model 4). This model of responses on the MACE was further supported by the exploratory factor analyses, where the 3-factor solution found a similar partitioning of the items. Notably, across samples, subscale scores derived from the peer victimization items were both reliable and interpretable insofar that the majority of their variance were attributable to the peer victimization factor even after accounting for the general maltreatment factor. It therefore may be valid to derive two score measures from the MACE, maltreatment by adults and by peers, but further research is needed to properly validate these measures.

Our results are consistent with a number of previous studies investigating the factor structure of other childhood maltreatment measures. For example, bifactor models of the Childhood Trauma Questionnaire have also revealed a dominant general maltreatment factor and weaker specific factors \citep{spinhoven2014childhood, hollerbach2018main, stagaki2022mediating}. Similar results were found in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}. Together, these and our current results underscore the challenge of developing versatile measures of childhood maltreatment. When the experience of distinct forms of abuse are correlated, it is difficult to measure any particular type of abuse independent of overall maltreatment.

>>>>

Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poor physical health \citep{wegman2009meta}. Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}. Thus, reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

@@

An overall shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance. One possibility is that the replication sample had simply experienced more adversity than the original sample. A second is that the differences in the demographic composition of the two samples (e.g. gender, age) explain the differences in endorsement rates. To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's endorsement response on an item is regressed against their rest score (i.e. their total score for all other items), sample membership, gender, and age.

@@

To evaluate the reliability of the MACE total and subscale scores, we calculated several model-based reliability indices using the factor loadings from the bifactor models. First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e. both general and specific factors). Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}). Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor. The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores. Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score. When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor \citep{rodriguez2016applying}. Values of $\omega_s$ approaching 1 indicate the specific factor, not the general factor, is the primary source of reliable variance in a subscale score. \cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable. In contrast, values of $\omega_s$ < 0.50 preclude the interpretation of subscale scores as primarily reflecting a specific factor \citep{gignac2013bifactor}.

To assess the unidimensionality of the MACE, we calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}). The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale. It thus measures the contribution of the general factor relative to the specific factors. As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble the item loadings that would be obtained by a one-dimensional model. ECV values greater than 0.70 often indicate a scale is essentially unidimensional \citep{rodriguez2016applying}. In turn, PUC quantifies how many inter-item correlations are accounted for only by the general factor. PUC is an important moderator of the ECV; as PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure. When PUC > 0.80, there is a low risk of bias when a multi- dimensional scale is treated as unidimensional \citep{reise2013multidimensionality}. We also calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading in a bifactor model, divided by the general factor loading from the bifactor model. RPB less than 10–15\% is acceptable \citep{muthen1987structural}.

Finally, we calculated the H index to assess the construct replicability of each factor \citep{hancock2001rethinking}. The H index is an estimate of how well a set of items represents a latent factor. H values closer to 1 indicate a well-defined latent factor with item loadings more likely to be stable across studies. In contrast, small H values suggest a poorly defined latent factor where item loadings are liable to change across studies. Because we are presently focused on evaluating the reliability of MACE total and subscale scores, we report the H values below but do not interpret them further.

@@

In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data from each sample separately. The goal of the EFA was to examine what factor structure would emerge with no \emph{a priori} restrictions placed on the data. We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models. We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (maladaptive parenting, peer victimization, reverse-scoring).

EFA was performed using the \textit{lavaan} (v0.6.11; \citealt{lavaan}) package available in R. Factors were extracted using the oblique geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings. We used two rotation criteria in order to examine the generalizability of the factor solutions across rotations. We again relied on traditional fit indices (i.e. RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory factor models to the response data of both samples.

@@

The item discrimination parameters of the unidimensional model were, on average, large for both the original sample (mean $\alpha$ = 1.357, 95\% HDI = 1.295 -- 1.414) and replication sample (mean $\alpha$ = 1.349, 95\% HDI = 1.272 -- 1.422). The standardized factor loadings were correspondingly large in magnitude. In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621). Most items had loadings greater than 0.50 (original: 80.0\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment. The rank correlation of general factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement between samples.

To begin characterizing the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric item correlation matrix for each sample. In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they were 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56. The first-to-second eigenvalue ratios of 3.84 and 3.02 are just above the 3-to-1 criterion often cited in determining whether item response data are essentially unidimensional \citep{embretson2013item}.

In sum, the fits of the one-factor models provide partial evidence in support of a unidimensional model of the MACE. Virtually all items exhibited moderate-to-large loadings onto the common factor. Moreover, the eigenvalues of the polychoric correlation matrix suggested one primary factor. However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be additional dimensions to the MACE. We turn next to the multidimensional confirmatory models to explore these possibilities.

@@

@@

Next we inspect the bifactor model with 10 specific factors, i.e. one per subscale on the MACE (model 2). In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864 with an average value of 0.575 (95\% HDI = 0.556 -- 0.593). In the replication sample, the general factor loadings ranged between 0.299 and 0.827 with an average value of 0.580 (95\% HDI = 0.560 -- 0.601). The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), suggesting there was little bias in the factor loadings of the one-factor models due to multidimensionality.

In contrast, the loadings on the specific factors were smaller and seldom exceeded loadings on the general factor. In the original dataset, the specific factor loadings ranged between 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341). In the replication dataset, the range was between 0.043 and 0.698, with an average value of 0.343 (95\% HDI = 0.315 -- 0.369). Only items from the peer verbal and physical abuse subscales had specific factor loadings of equal or greater magnitude to those on the general factor.

@@

We found that the MACE total score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965). The $\omega_h$ value was 0.924 for both samples indicating that 96.0\% and 95.8\% of the reliable variance in the total score is attributable to general maltreatment. (This result is not very surprising given the large number of subscales composed of few items.) Importantly, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively. In conjunction with a large PUC value of 0.912, these results provide strong evidence for a dominant general maltreatment factor in the MACE.

The reliability of the MACE subscale scores were more variable in comparison. The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample, with smaller $\omega$ values for subscales with fewer items. Crucially, the $\omega_s$ values were small for both the original sample (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and replication sample (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565). That is, the MACE subscale scores contain almost no reliable variance after controlling for the general maltreatment factor. They should not therefore be interpreted as reflecting specific forms of maltreatment. The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores was independent of general maltreatment (original = 70.7\%; replication = 66.2\%).

The results of the full bifactor model were clear: responses on the MACE are explained by a dominant general maltreatment factor characterized by large loadings from the majority of items. In contrast, specific factors corresponding to each MACE subscale are weak and characterized mostly by small factor loadings. Whereas the MACE total score is a reliable and valid measure of overall maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to particular types of abuse. The question remains, however, if there are other subscale scores to be derived from the MACE.

\subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}

We turn our attention next to the bifactor S-1 model with one specific factor (neglect; model 3). The loadings on the general factor largely resembled those of the bifactor model. In the original sample, the loadings on the general maltreatment factor ranged between 0.150 and 0.864 with an average value of 0.577 (95\% HDI = 0.558 -- 0.594). In the replication sample, the general factor loadings ranged between 0.395 and 0.831 with an average value of 0.580 (95\% HDI = 0.558 -- 0.601).

The loadings on the neglect specific factor were smaller on average. In the original sample, the average specific loading was 0.410 (95\% HDI = 0.373 -- 0.454); in the replication sample, it was 0.454 (95\% HDI = 0.409 -- 0.499). The specific loadings displayed a troubling pattern where only the six reverse-scored items exhibited moderate loadings on the specific factor; the remaining four item exhibited negligible factor loadings. That is, the ``neglect'' factor appears instead to reflect a reverse-scoring methods factor.

Moving on to the reliability measures, we again found that the total score exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959). Moreover, the $\omega_h$ values indicated the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922). The reliability of the ``neglect'' subscale score was also good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921). The $\omega_s$ values, however, indicated the neglect subscale score contained less than half reliable variance after controlling for the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).

In summary, the results do not support the use of the MACE to make separate threat and deprivation subscale scores. The neglect subscale scores posses only a minority of reliable variance unique from general maltreatment thereby precluding meaningful interpretation of these scores. Furthermore, the majority of items measuring neglect on the MACE seem to be contaminated by a reverse-scoring methods factor.

@@

Finally, we inspect the bifactor S-1 model with two specific factors measuring peer victimization and reverse-scoring (model 4). In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864 with an average value of 0.563 (95\% HDI = 0.545 -- 0.580). In the replication sample, the general factor loadings ranged between 0.298 and 0.852 with an average value of 0.572 (95\% HDI = 0.550 -- 0.593). Interestingly, the loadings on the specific factors were of approximately equal magnitude. In the original dataset, the specific factor loadings ranged 0.327 to 0.789 with an average value of 0.550 (95\% HDI = 0.522 -- 0.576). In the replication dataset, the range was between 0.183 and 0.684 with an average value of 0.526 (95\% HDI = 0.492 -- 0.556).

As with the other bifactor models, the total score under this model exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961). Again the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.896; replication: $\omega_h$ = 0.904). The peer victimization subscale score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868). The corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample. That is, slightly more than half of the variance in peer victimization scores (original: 64.0\%; replication: 56.8\%) is independent of the general maltreatment factor. These results suggest the presence a secondary peer victimization factor in the MACE (albeit weakly so) in addition to a primary general maltreatment factor.

Briefly, we note that the subscale scores formed by the reverse-scored items were also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915). Moreover, the corresponding $\omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498). In other words, these scores are also have majority reliable variance independent of general maltreatment. It is difficult to interpret these scores, however, as it is unclear what they primarily reflect: neglect experiences, wording effects, and/or scoring-induced methods artifact.

@@

The goodness-of-fit measures for the exploratory factor models to the response data by sample are summarized in Table \ref{tab:efa_diagnostics}. The fit indices indicated that all exploratory models provided an adequate fit to the data. The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are presented in Figure \ref{fig:efa_cf}. (The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, we observed a primary maladaptive parenting factor and a secondary peer victimization factor. In contrast, the replication sample produced a primary maladaptive parenting/peer maltreatment factor with a secondary reverse-scoring factor. Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.

The 3-factor solution for each sample were similar. EFA produced a primary maladaptive parenting factor with secondary peer victimization and reverse-scoring factors. (Interestingly, items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, raising some questions about the replicability of the factor structures.) To conclude, the structure of both the 2- and 3-factor exploratory factor models resembled the bifactor S-1 model with two specific factors (model 4), providing further support for that model of the factor structure of the MACE.

@@

The MACE is a promising measure of childhood maltreatment \citep{georgieva2022systematic}. Despite its increasing popularity, there has been little investigation into the MACE's structural validity \citep{saini2019systematic} leaving unanswered questions about its factor structure and the validity of its total/subscale scores. Here we investigated the factor structure of the MACE, using both confirmatory and exploratory item response modeling, in two independent samples. Bifactor modeling revealed a dominant general maltreatment factor, characterized by uniformly strong loadings across items, and much weaker specific factors. The general factor explained approximately 70\% of the reliable variance in the response data across samples. Moreover, the general factor loadings from the bifactor models largely resembled the corresponding loadings from a unidimensonal model. Together, these results support the conclusion that the MACE is essentially unidimensional.

We also calculated reliability indices for the MACE total and subscale scores. We found that the MACE total score has excellent reliability and is a valid measure of overall maltreatment. In contrast, we found that virtually no MACE subscale scores were reliable measures of particular types of maltreatment. For all but one of the 10 MACE subscales, the majority of reliable variance in scores was attributable to the general factor. The same was true of deprivation subscale scores, which were also contaminated by a reverse-scoring methods factor. In short, our results further validate the use of the MACE total score, but caution against the use of MACE subscale scores. Indeed, MACE subscale scores seldom reflected their intended construct but instead reflected general maltreatment.

Of the four confirmatory factor models we tested, the most parsimonious and best- fitting was a bifactor S-1 model with a prominent general maltreatment factor and two specific factors measuring peer victimization and reverse-scoring. This model of the MACE factor structure was additionally supported by the exploratory factor analyses, which produced a qualitatively similar partitioning of the items. Thus, it appears the MACE is essentially unidimensional with secondary factors affecting a small subset of items. Interestingly, we found the peer victimization subscale scores to be reliable; a majority of their variance was attributable to the peer victimization factor even after controlling for general maltreatment. It may therefore be appropriate to produce a peer victimization subscale score from the MACE, but additional research is needed to validate this measure.

Our results are consistent with a number of previous studies investigating the factor structure of other childhood maltreatment questionnaires. For example, bifactor models of the Childhood Trauma Questionnaire have also revealed a dominant general maltreatment factor and weaker specific factors \citep{spinhoven2014childhood, stagaki2022mediating}. Similar results were found in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}. Together, these and our current results underscore a challenge in investigating the consequences of childhood maltreatment. When maltreatment experiences tend to co-occur, it is difficult to measure the effects of a particular type of adversity unique from overall maltreatment.

Our results are also worth considering alongside previous studies that used MACE subscale scores to investigate the link between particular types of childhood maltreatment and assorted psychobiological outcomes. For example, researchers have looked at the relationship between MACE subscale scores and risk of depression \citep{gerke2018childhood}, dissociative symptoms, \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}. Still others have used threat and deprivation subscale scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}. Our results are clear that these scores chiefly reflect general maltreatment and not particular types of adversity, which presents a challenge for interpreting the conclusions of those studies. It is worth noting that some of these studies attempted to control for general maltreatment by entering all of the MACE subscale scores simultaneously into sophisticated multivariate analyses (e.g. random forest regression). This approach is unsatisfactory for two reasons. First, random forest models can produce spurious results when dealing with collinear predictors \citep{gregorutti2017correlation}. Second, even if one were to properly control for general maltreatment, our results indicate that the remaining reliable variance in most subscale scores is so low as to cast doubt about the utility of these inferences.
